---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load the data:
```{r load}
activity <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?

```{r histogram}
## calculate total number of steps each day for use in the histogram and the calculations of mean and median after
totalSteps <- with (activity, tapply(steps, date, sum))

## Histogram of the total number of steps taken each day
hist(totalSteps, breaks=20, col="red", main="Histogram of the total number of steps taken each day", xlab="Number of steps")
```

```{r calculate}
meanTotalSteps <- round(mean(totalSteps[!is.na(totalSteps)]), 2)
medianTotalSteps <- median(totalSteps[!is.na(totalSteps)])
```

Mean total number of steps taken per day:       `r meanTotalSteps`  
Median total number of steps taken per day:     `r medianTotalSteps`

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
