---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load the data:
```{r load}
activity <- read.csv("activity.csv")
```



## What is mean total number of steps taken per day?

```{r histogram}
## calculate total number of steps each day for use in the histogram and the calculations of mean and median after
totalSteps <- with (activity, tapply(steps, date, sum))

## Histogram of the total number of steps taken each day
hist(totalSteps, breaks=20, col="red", main="Histogram of the total number of steps taken each day", xlab="Number of steps")
```

```{r calculate}
meanTotalSteps <- round(mean(totalSteps[!is.na(totalSteps)]), 2)
medianTotalSteps <- median(totalSteps[!is.na(totalSteps)])
```

Mean total number of steps taken per day:       `r meanTotalSteps`  
Median total number of steps taken per day:     `r medianTotalSteps`



## What is the average daily activity pattern?
```{r time}
meanIntervalSteps<-with (activity, tapply(steps[!is.na(steps)], interval[!is.na(steps)], mean))
intervals<-unique(activity$interval)
plot(intervals,meanIntervalSteps, type="l", main="Time series plot of average steps for each 5 minute interval", xlab="Time intervals", ylab="Average number of steps")
```

```{r interval}
maxInterval <- intervals[match(max(meanIntervalSteps),meanIntervalSteps)]
```

On average across all the days in the dataset, the maximum number of steps is in interval:      `r maxInterval`



## Imputing missing values
```{r}
totalMissing <- length(activity$steps[is.na(activity$steps)])
```

The total number of missing values in the dataset is:   `r totalMissing`


The strategy chosen for filling the missing values in the dataset is to replace NAs with the mean for the corresponding 5-minute interval

```{r filled}
fill <- function(x1, x2){
        for (i in 1:length(x1)){
                j=i %% 288
                if(is.na(x1[i])){
                        if(j==0){j=288}
                        x1[i]=x2[j]
                }
        }
        x1
}

filledSteps <- activity$steps
filledSteps <- fill(filledSteps, meanIntervalSteps)

filledActivity <-data.frame(filledSteps, activity$date, activity$interval)
```

```{r histogram2}
## calculate total number of steps each day with filled dataset
totalSteps2 <- with (filledActivity, tapply(filledSteps, activity.date, sum))

## Histogram of the total number of steps taken each day
hist(totalSteps2, breaks=20, col="red", main="Histogram of the total number of steps (with substitution)", xlab="Number of steps")
```

```{r calculate2}
meanFilledTotalSteps <- round(mean(totalSteps2, 2))
medianFilledTotalSteps <- median(totalSteps2)

meanDiff <- meanFilledTotalSteps - meanTotalSteps
medianDiff <- round(medianFilledTotalSteps - medianTotalSteps, 2)
```

Mean total number of steps taken per day (with substitution):       `r meanFilledTotalSteps`  
Median total number of steps taken per day (with substitution:     `r medianFilledTotalSteps`  


Difference in means for the two datasets:       `r meanDiff`  
Difference in medians for the two datasets:     `r medianDiff`  

So the impact of imputing missing data on the estimates of the total daily number of steps is very small when using the mean for the corresponding 5-minute interval as a substitute



## Are there differences in activity patterns between weekdays and weekends?
